### 2018-03-20 Paper/jaccardbaetica.tex
updated introduction

% Debate

@@

The good condition of the Baetican lands stimulated by the high demand of Roman provinces allowed to develop a massive infrastructure of olive oil production. This product was distributed in large amount of amphorae along the province, mostly to supply the Roman Army (citar Carreras o Blazquez illo).

The production and distribution of olive oil in this ancient province were growing exponentially during almost three centuries (cronologiaaa CITAR). As a results, hundred of amphora workshops were created to support this high demand. These workshops were located along the river Guadalquivir and its tributary Genil supplying from the riverine connectivity to the Mediterranean and Atlantic (CITAR GARCIA VARGAS)

@@

%Hyphotesis
However, it remains under debate how this province was organized and whether it is possible to identify patterns in the olive oil trade.

%hablar un poco del comercio del aceite de oliva quizás citando el paper de Williams, Peacok y mattlingly el colega
%hablar de las tres áreas de producción: hispalis, corduba y astigi

@@

Aquí poner algo sobre esto \citep{remesal_anforas_2004}

Here, therefore, this study aim to explore the effect of Baetican olive oil production in the Roman Empire
of production organization over Baetican olive oil production.

@@

Our case study examines the relation between the distribution of amphoric stamps and the workshops. We studied a dataset of 3787 stamps collected from different amphora workshops in Baetica provice (see CEIPAC database). The workshops were situated in different locations in Baetica province, along the river Guadalquivir and its tributary Genil in order to detect similarities between stamps from workshops and spatial distance.

@@

The chronology in the workshops is widely diverse from the first to the third centuries AD \citep{garcia_vargas_enrique_formal_2010}. However, some stamps show a more specific chronology while the majority of them display a large activity of production being difficult to specify an accurate chronology. A reason may be that most of them were partially excavated and only focused on archaeological surveys to collect the maximum stamps as possible (CITAR)

>>>>

% Debate
In archaeology, this debate has been widely studied in the last centuries (CITAR). Specifically, our work pretend to identify production processes focus on the olive oil trade.

@@

The high demand of Roman provinces stimulated by the good condition of the Baetica lands allowed to develop a massive infrastructure of olive oil production. This product was distributed in large amount of amphorae along the province, mostly to supply the Roman Army and Italy \citep{blazquez_exportacion_1980}.

The production and distribution of olive oil in this ancient province were growing exponentially during almost three centuries (cronologiaaa CITAR). As a results, hundred of amphora workshops were created to support this high demand. These workshops were located along the river Guadalquivir and its tributary Genil supplying from the riverine connectivity to the Mediterranean and Atlantic \citep{garcia_vargas_enrique_formal_2010}.

@@

%Hyphotesis
However, some challenges remain under discussion such as how this province was organized and whether it is possible to identify production patterns in the olive oil trade (AMPLIAR Y METER SELLOS). Neither written evidences have been found that it could explain the economic role of Baetica province in the Roman organization. On the other hand, archaeological evidence show a highly specialized production with a long activity in a this specific area with apparently few changes \citep{remesal_anforas_2004}.

%hipotesis: si esto es así entonces es asá

Here, therefore, this study propose an approach to explore the effect of Baetican olive oil production by estimating of similarity index (COPIADO CAMBIAR). The assumption of this analysis is that closer workshops concentrate similar amphoric stamps in a specific area than the farthest workshops. By contrast, if we found similar amphoric stamps in different areas between workshops then the correlation between spatial distance would not be valued.

%resumen del paper

%hablar de las tres áreas de producción: hispalis, corduba y astigi

@@

Aquí poner algo sobre esto

@@

Our case study examines the relation between the distribution of amphoric stamps and the workshops. We studied a dataset of 3787 stamps collected from different amphora workshops in Baetica provice by CEIPAC database \citep{remesal_centro_2015} (see CEIPAC database). The workshops were situated in different locations in Baetica province, along the river Guadalquivir and its tributary Genil in order to detect similarities between stamps from workshops and spatial distance.

@@

The chronology in the workshops is widely diverse from the first to the third centuries AD (CITAR). However, some stamps show a more specific chronology while the majority of them display a large activity of production being difficult to specify an accurate chronology. A reason may be that most of them were partially excavated and only focused on archaeological surveys to collect the maximum stamps as possible (CITAR)
